Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-07 talk/pepm2011/paper.tex
this won't fit
these are all done
kill more XXXs

\authorinfo{Carl Friedrich Bolz \and Antonio Cuni \and Maciej Fijałkowski \and Michael Leuschel \and Samuele Pedroni \and Armin Rigo}
           {Heinrich-Heine-Universität Düsseldorf, STUPS Group, Germany XXX}
           {cfbolz@gmx.de, anto.cuni@gmail.com, fijal@merlinux.eu,
           samuele.pedroni@gmail.com, arigo@tunes.org}

@@

XXX drop the word "allocation removal" somewhere

XXX define "escape analysis"

@@

XXX need to re-target introduction a bit to fit PEPMs focus

@@

If the function is executed using the tracing JIT, with \texttt{y} being a
\texttt{BoxedInteger}, the produced trace looks like
Figure~\ref{fig:unopt-trace}. The operations in the trace are shown indented to
correspond to the stack level of the function that contains the traced
operation. The trace also shows the inefficiencies of \texttt{f} clearly, if one
looks at the number of \texttt{new} (corresponding to object creation),
\texttt{set/get} (corresponding to attribute reads/writes) and
\texttt{guard\_class} operations (corresponding to method calls).

@@

XXX simplify traces a bit more

In the next section, we will see how this can be improved upon, using partial
evaluation. XXX

@@

\footnote{This section is a bit
science-fictiony. The algorithm that PyPy currently uses is significantly more
complex and much harder than the one that is described here. The resulting
behaviour is very similar, however, so we will use the simpler version (and we
might switch to that at some point in the actual implementation).}

@@

The reason why we cannot optimize the remaining allocations away is because
their lifetime crosses the jump. To improve the situation, a little trick is
needed. The trace in Figure~\ref{fig:step1} represents a loop, i.e. the jump at the end jumps to
the beginning. Where in the loop the jump occurs is arbitrary, since the loop
can only be left via failing guards anyway. Therefore it does not change the
semantics of the loop to put the jump at another point into the trace and we
can move the \texttt{jump} operation just above the allocation of the objects that
appear in the current \texttt{jump}. This needs some care, because the arguments to
\texttt{jump} are all currently live variables, thus they need to be adapted.

If we do that for our example trace, the trace looks like in Figure~\ref{fig:step2}.

@@

Now the lifetime of the remaining allocations no longer crosses the jump, and
we can run our partial evaluation a second time, to get the trace in
Figure~\ref{fig:step3}.

@@

% subsection Virtualizables (end)
XXX store sinking

do we need this?

>>>>

\authorinfo{Carl Friedrich Bolz \and Antonio Cuni \and Maciej Fijałkowski \and Michael Leuschel \and Samuele Pedroni \and Armin Rigo}
           {Heinrich-Heine-Universität Düsseldorf, STUPS Group, Germany XXX}
           {cfbolz@gmx.de, anto.cuni@gmail.com, fijal@merlinux.eu,
           leuschel@cs.uni-duesseldorf.de, samuele.pedroni@gmail.com, arigo@tunes.org}

@@

@@

@@

If the function is executed using the tracing JIT, with \texttt{y} being a
\texttt{BoxedInteger}, the produced trace looks like
Figure~\ref{fig:unopt-trace}. The operations in the trace are shown indented to
correspond to the stack level of the function that contains the traced
operation. The trace also shows the inefficiencies of \texttt{f} clearly, if one
looks at the number of \texttt{new} (corresponding to object creation),
\texttt{set/get} (corresponding to attribute reads/writes) and
\texttt{guard\_class} operations (corresponding to method calls).
In the rest of the paper we will see how this trace can be optimized using
partial evaluation.

@@

@@

@@

The reason why we cannot optimize the remaining allocations away is because
their lifetime crosses the jump. To improve the situation, a little trick is
needed.\footnote{The algorithm that PyPy currently uses is significantly more complex
than the one that is described here. The resulting behaviour is nearly
identical, however, so we will use the simpler version (and plan to switch to
that at some point in the actual implementation).}
The trace in Figure~\ref{fig:step1} represents a loop, i.e. the jump at the end jumps to
the beginning. Where in the loop the jump occurs is arbitrary, since the loop
can only be left via failing guards anyway. Therefore it does not change the
semantics of the loop to put the jump at another point into the trace and we
can move the \texttt{jump} operation just above the allocation of the objects that
appear in the current \texttt{jump}. This needs some care, because the arguments to
\texttt{jump} are all currently live variables, thus they need to be adapted.

@@

If we do that for our example trace, the trace looks like in Figure~\ref{fig:step2}.
Now the lifetime of the remaining allocations no longer crosses the jump, and
we can run our partial evaluation a second time, to get the trace in
Figure~\ref{fig:step3}.

@@

% subsection Virtualizables (end)
